### 2011-12-22 papers/PLDI2012/response_PLDI2012.txt
review + brouillon rebuttal

"the title is fairly vague and non optimal -- perhaps
"succinct" should be replaced by "disjunctive"?"
Path Focusing technique and our combined technique (with or without disjunctive
invariants) do not work on the classical control flow graph, but on a multigraph
where all the paths inside loops are distinguished. The term "succint" refers to
the fact we actually do not compute all the paths of this multigraph, but work
on an succint representation thereof (using an SMT formula), thus avoiding the
enumeration of an exponential number of paths in a huge number of cases.

"The results are unsurprising"
The experimental results show that techniques that distinguish paths inside a
program give more precise results, but, more importantly, at a reasonable cost
(which was not evident).
It is also interesting to see that more complex technique perform sometimes
faster that more naive approaches.

"The problem with path focusing is that (1) it will process all
executable paths at once, including potentially many redundant paths
that don't contribute to the solution; and (2) its widening operation
lacks the precision conferred by guided static analysis."

This is not exactly the point. The problem with path focusing is that it will compute
the fixpoint iterations on the whole graph at once, using classical widening
operations in order to guarantee the convergence.
When the fixpoint has been reached, it finally apply some narrowing iterations
to recover precision.
The idea here is to apply some narrowing iterations during the analysis, in
order to recover some precision before continuing.
To clarify, widening is the same, but we perform regularly narrowing iterations,
and not only at the end of the ascending iterations.

"how much benefit this technique confers to other types of abstract
domains, including those that don't require widening"
Intuitively, Guided Static Analysis technique performs better because it
recovers precision lost by widening. It will probably not increase the precision
when using domains without widening/narrowing. However, the iteration strategy
will be changed, so the computation time may differ compared to a classical
abstract interpretation. We can do the same remarks on our combined technique
compared to Path Focusing.

"I'd like to know more about how applicable the technique
is to other kinds of abstract domains"
The general idea does not depend on the abstract domain. However, we have to
transform an abstract value into an SMT formula. Such transformation is easy in
our case, since it is a cunjunction of linear inequalities, but may still
require some research for other abstract domains that could require other
theories (for instance, arrays could be
encoded in SMT using uninterpreted functions, etc.)

"2 to 5% difference, does it make a difference at all?"
Our prototype of analyzer performs many over-approximations (no memory model,
etc.), thus leading to invariants that are often very unprecise.
Intuitively, I would say we could have obtained more difference between the
different techniques by reducing the number of over-approximation we do for the
time being.