### 2012-05-23 papers/SAS2012/implicitization_article.tex
mise en page version finale, debut

\documentclass[a4paper]{amsart}

\usepackage{ifxetex}
\ifxetex
\else
\usepackage[utf8]{inputenc}
\fi

\usepackage{amsmath}
\usepackage{mathcomp,amsfonts,amssymb}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}
\usepackage{paralist}
\usepackage{graphicx}
%\usepackage{color}
\usepackage{placeins}

\usepackage{enumitem}
\usepackage{dcolumn}

\usepackage[numbers]{dmnatbib}

\usepackage{listings}
\lstset{language=C,mathescape=true,numbers=left,numbersep=0.8ex,tabsize=2,breaklines=true,basicstyle={\rm\fontfamily{phv}\small\selectfont}}

\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}

\tikzstyle{state}=[circle,fill=black!25,minimum size=13pt,inner sep=0pt]
\tikzstyle{rstate}=[rectangle,fill=black!25,minimum size=13pt,inner sep=0pt]
\tikzstyle{transition}=[rectangle,semithick,draw=black!75,
  			  minimum size=4mm]
\tikzstyle{transition2}=[transition,rectangle,thick,dashed,
  			  minimum size=4mm]
\tikzstyle{PRstate}=[circle,double,draw,fill=blue!15,minimum size=13pt,inner sep=0pt]
\tikzstyle{polyhedra}=[blue!25,opacity=0.5,pattern=north west lines,pattern
color=blue]
\tikzstyle{line}=[black,thick]

\newcommand{\mytitle}{Succinct Representations for Abstract Interpretation}
\usepackage[pdftitle={\mytitle},pdfauthor={Julien Henry, David Monniaux, Matthieu Moy}]{hyperref}

\newcommand{\abstr}[1]{#1^\sharp}
\newcommand{\parts}[1]{\mathscr{P}(#1)}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\widening}{\mathop{\triangledown}}

\graphicspath{{figures/}{gnuplot/}}

% commentaires dans les algorithmes crees avec le package algorithmic
\renewcommand{\algorithmiccomment}[1]{ {\color{gray} // #1}}

\usepackage{ifthen}
\newboolean{draft}
%\setboolean{draft}{true}
\setboolean{draft}{false}

\ifthenelse{\boolean{draft}}{
  \newcommand{\avirer}[1]{{\color{red} #1}}
  \newcommand{\MM}[1]{{\color{blue} TODO(MM): #1}}
  \newcommand{\JH}[1]{{\color{teal} TODO(JH): #1}}
  \newcommand{\DM}[1]{{\color{violet} TODO(DM): #1}}
}{
  \newcommand{\avirer}[1]{}
  \newcommand{\MM}[1]{}
  \newcommand{\JH}[1]{}
  \newcommand{\DM}[1]{}
}

\renewcommand{\baselinestretch}{1}

\begin{document}
\title[Succinct representations for abstract interpretation]{Succinct representations for abstract interpretation \\Combined analysis algorithms and experimental evaluation}
%\thanks{\raisebox{-0.5em}{\includegraphics[height=1.5em]{ASOPT_officiel_nb}} This work was partially funded by ANR project ``ASOPT''}

\author{Julien Henry}
\address{J. Henry\\VERIMAG\\2 av de Vignate\\38610 Gières\\France}
\email{Julien.Henry@imag.fr}
\urladdr{http://www-verimag.imag.fr/~jhenry/}
\thanks{Julien Henry is a graduate student at Université Joseph Fourier, VERIMAG laboratory. VERIMAG is a joint laboratory of Université Joseph Fourier, CNRS and Grenoble-INP}

\author{David Monniaux}
\address{D. Monniaux\\VERIMAG\\2 av de Vignate\\38610 Gières\\France}
\email{David.Monniaux@imag.fr}
\urladdr{http://www-verimag.imag.fr/~monniaux/}
\thanks{David Monniaux is researcher at CNRS, VERIMAG laboratory}

\author{Matthieu Moy}
\address{M. Moy\\VERIMAG\\2 av de Vignate\\38610 Gières\\France}
\email{Matthieu.Moy@grenoble-inp.fr}
\urladdr{http://www-verimag.imag.fr/~moy/}
\thanks{Matthieu Moy is assistant professor at Grenoble-INP, VERIMAG laboratory}

\begin{abstract}
\input{abstract}
\end{abstract}

\maketitle

\input{implicitization}

\FloatBarrier

\bibliographystyle{dmplainnat}
\bibliography{implicitization}
\end{document}